Formation of porous crystals via viscoelastic phase separation

### 2016-12-20 nature/RebuttalLetter.tex

In the new version of the manuscript we have tried to incorporate all the Referee's previous suggestions. 
Following the suggestion, we have explicitly mentioned the work of Poon and coworkers in connection to the Bergeron and Ostwald ripening processes.  
We have also added the two recommended references by Sanz et al. to the new version of the manuscript. 
On the doubt of the referee concerning the novelty of our work, we hope that our new analysis explained above provides a 
new physical insight on the physical condition required for the formation of porous crystal gels besides its new kinetic route. 
We also simplified the title to avoid misunderstanding of our message.

Regarding our experimental protocol, it is indeed reminiscent of salt injection, but differs in two important ways. (i) Rather than an injection salts are introduced through a membrane. This dialysis protocol allows salt introduction without external flow. (ii) We are in a regime where, upon salt introduction, charges are screened throughout the sample almost instantaneously, that is to say in a few Brownian times. This is due to the short length the ions have to diffuse from the reservoir to the top of the sample and to the large diffusion constant ratio between the ions and our large colloidal particles.
We believe this to be the first time this cell design has been applied for confocal microscopy observation at a single-particle level and used to study the formation of porous crystals in gels, which is the key to access the early stage of viscoelastic phase separation at a single particle resolution with full particle tracking. 
For example, the new analysis shown in new Fig.~2 is only possible with this new experimental protocol and full particle tracking.  
Finally, the last sentence of the abstract has been modified, following the kind suggestion of the Referee.

@@

%\textcolor{red}{The Referee insists on having surface particles represented in Fig.6b. Even if I think that the figure becomes more complicated this way, maybe we should just update the figure. I'm sending Mathieu the dataset with the fraction of surface particles and liquid particles in case you agree on this change. If you do, I'll also update the description in the manuscript.}

%\textcolor{red}{We remind that the 54/34/11 abundances was obtained by looking at the phase history of each colloidal particle. Fig. 6b and Supplementary Fig. 4 don't contain information about the phase history of individual particles, just about the total number of particles in each phase, and the time-dependence of the transition probabilities. Of course these quantities are closely connected, and we hope to have provided a detailed description of the process (??).}

\vskip 5mm

We hope that the Referee will find that the new analysis has revealed a new mechanism behind 
crystal-gel formation and increased the novelty of our manuscript considerably and 
the revised manuscript is now suitable for publication in Nature Materials.

### 2016-12-20 crystal-gel/nature/crystal_gel_r.tex
rephrase crystal growth implying less surface

In Fig.~\ref{fig:transitions}b, we show  the fraction of particles in each different phase for the crystallization trajectory at
$\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
\textcolor{blue}{Surface, gas and liquid particles are identified from the subset of non-crystalline particles depending
on their neighbouring particles. A particle with at least two neighbouring crystal particles is classified as
\emph{surface}. A particle that is neither solid nor surface, and has at least four neighbouring particles
is classified as \emph{liquid}. A particle which is neither solid, surface nor liquid is classified as \emph{gas}.}
The process of crystallization is characterized 
by a steep decrease in liquid particles, as they transform into small crystalline nuclei inside the liquid domains. This decrease
is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. \textcolor{blue}{The crystalline nuclei are initially composed mostly of surface particles, but as the growth of nuclei becomes more compact the number of surface particles slowly decreases, while bulk crystals keep increasing}.
After the onset of a steady-state gas population, there are three growth routes for the crystal. 
Direct crystallization is the process by which crystals grow by incorporating nearby liquid particles. 
In the Bergeron process, liquid droplets first evaporate and the resulting gas phase contributes to the growth of crystalline regions. Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal particles are transferred to larger nuclei.
\textcolor{blue}{Since we have access to individual particle trajectories, we can directly assess the relative importance of these three growth channels.}
For every time frame, we assign each particle a state between gas, liquid and crystal. In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_{\rm B}$. We then measure the fraction of trajectories with different transition histories.
The different crystallization routes are depicted in the diagram of Fig.~\ref{fig:transitions}c.
Direct crystallization accounts for 54\% of particles trajectories in which gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. 
The Bergeron process accounts for 35\% of particle trajectories in which liquid particles transition to the gas state before crystallizing. 
Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning to the crystal state. \textcolor{blue}{Here we note that the Bergeron and Ostwald ripening processes both take place only when the crystallites are 
surrounded by the gas phase due to the non-existence of a liquid-crystal coexistence. Such a gas-crystal coexistence was speculated  by Poon and his coworkers on the basis of free-energy argument \cite{poon1999cpm,renth2001phase}.}

>>>>

In Fig.~\ref{fig:transitions}b, we show  the fraction of particles in each different phase for the crystallization trajectory at
$\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
\textcolor{blue}{Surface, gas and liquid particles are identified from the subset of non-crystalline particles depending
on their neighbouring particles. A particle with at least two neighbouring crystal particles is classified as
\emph{surface}. A particle that is neither solid nor surface, and has at least four neighbouring particles
is classified as \emph{liquid}. A particle which is neither solid, surface nor liquid is classified as \emph{gas}.}
The process of crystallization is characterized 
by a steep decrease in liquid particles, as they transform into small crystalline nuclei inside the liquid domains. This decrease
is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. \textcolor{blue}{The crystalline nuclei are initially composed mostly of surface particles, but as the nuclei grow they becomes more compact and merge together such that the number of surface particles slowly decreases, while bulk crystals keep increasing}.
After the onset of a steady-state gas population, there are three growth routes for the crystal. 
Direct crystallization is the process by which crystals grow by incorporating nearby liquid particles. 
In the Bergeron process, liquid droplets first evaporate and the resulting gas phase contributes to the growth of crystalline regions. Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal particles are transferred to larger nuclei.
\textcolor{blue}{Since we have access to individual particle trajectories, we can directly assess the relative importance of these three growth channels.}
For every time frame, we assign each particle a state between gas, liquid and crystal. In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_{\rm B}$. We then measure the fraction of trajectories with different transition histories.
The different crystallization routes are depicted in the diagram of Fig.~\ref{fig:transitions}c.
Direct crystallization accounts for 54\% of particles trajectories in which gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. 
The Bergeron process accounts for 35\% of particle trajectories in which liquid particles transition to the gas state before crystallizing. 
Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning to the crystal state. \textcolor{blue}{Here we note that the Bergeron and Ostwald ripening processes both take place only when the crystallites are 
surrounded by the gas phase due to the non-existence of a liquid-crystal coexistence. Such a gas-crystal coexistence was speculated  by Poon and his coworkers on the basis of free-energy argument \cite{poon1999cpm,renth2001phase}.}
